---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---


## Setup

### Load packages
for this project I will be useing ggplot and ggmosaic to display the data and dplyr for our dataframe manipulation

```{r load-packages, message = FALSE}
library(ggplot2)
library(ggmosaic)
library(dplyr)
```

### Load data


```{r load-data}
load("brfss2013.RData")
```


* * *

## Part 1: Data

The Behavioral Risk Factor Surveillance System or BRFSS for short is a health survey originating in 1984 with 15 states conducting monthly telephone interviews who’s partipicents are non-institutionalized adults (18 and over). According to the [BRFSS](https://www.cdc.gov/brfss/annual_data/2013/pdf/Overview_2013.pdf) during telephone interviews the subject were questioned about there behavioral risk factors including but not limited to “tobacco use, HIV/AIDS knowledge and prevention, exercise, immunization, health status, healthy days — health-related quality of life, health care access, inadequate sleep, hypertension awareness, cholesterol awareness, chronic health conditions, alcohol consumption, fruits and vegetables consumption, arthritis burden, and seatbelt use”.  Today the survey covers all 50 states, the District of Columbia, Guam, and Puerto Rico “making it the larges continuously conducted heath survey system in the world … completing more than 400,000 interviews each year.” 

The BRFSS land line and recently cell phone, interview data has been randomized by each states collection agency.  While “No direct method of accounting fort non-telephone coverage is employed by the BRFSS”, the BRFSS uses an “iterative proportional filling (or “ranking”)’ To account for the households with out phone access.  This methodology weights the survey data to underlying demographics so that under represented groups are better represented in the final data set. “The use of ranking has been shown by researches to reduce error within estimates.”

The inability of the researchers to question everyone that they have randomly sampled and the fact that during the interview the participant has the option to terminate the interview (resulting in many unanswered questions). The BRFSS while a great tool for identifying correlation can not used in causal analysis.  What ever conclusions, will need further independent study with a properly randomized sample study. 

STILL NEED TO TALK ABOUT THE POPULATION STUDIED

* * *


## Part 2: Research questions

```{r echo=FALSE, results="asis", tidy=FALSE}
cat("[**Research quesion 1:**](#ResearchQuestion1)")
```

Explore the correlation between the number or fruits and the number of vegetables the participants eat each month. My assumption is that fruit and vegetable consumption is positively correlated.  After this is complete check if the number of fruits and vegetables has a positive correlation with the participants over all health.

```{r echo=FALSE, results="asis", tidy=FALSE}
cat("[**Research quesion 2:**](#ResearchQuestion2)")
```
Explore the relationship between smoking and educations. How predictive is education level to if they smoke or not

```{r echo=FALSE, results="asis", tidy=FALSE}
cat("[**Research quesion 3:**](#ResearchQuestion3)")
```
Question 3


* * *

## Part 3: Exploratory data analysis

##### **Research quesion 1:** {#ResearchQuestion1}

*Find the index values for the columns I wish to analysis*
```{r echo=TRUE}
grep("vegetab1", colnames(brfss2013))
grep("fruit1", colnames(brfss2013))
```
*Create a data frame with just the vegetables and fruit columns*
```{r}
cols = c(82, 78)
fruits_and_vegetables <- brfss2013[,cols, drop=FALSE]
```
*The data is stored in the database as such: when the value is [101-199]  the tens and units columns gives you the integer amount per day, when the value is [201 - 299] the tens and unit columns gives you the integer amount per week, when the value is [301 - 300] the tens and units give you the integer amount per month, 300 means less than once per month and 0 means never. (for our purpose we will treat 300 and 0 as 0)  Next create columns with mutated values that represent the number of monthly fruits and vegetables eaten, approximate the number of weekly fruits and vegetables by multiplying the value by 4 (4 weeks a month), and approximate the number of daily fruits and vegetables by multiplying the daily value by 30 (30 days in a month)*
```{r}
fruits_and_vegetables <- fruits_and_vegetables %>%
  mutate(veg_month = ifelse(vegetab1<=399 & vegetab1 >=301, (vegetab1 - 300), ifelse(vegetab1<=299 & vegetab1>=201,(vegetab1-200)*4, ifelse(vegetab1<=199 & vegetab1>=101,(vegetab1-100)*30, 0)))) %>%
   mutate(fruit_month = ifelse(fruit1<=399 & fruit1 >=301, (fruit1 - 300), ifelse(fruit1<=299 & fruit1>=201,(fruit1-200)*4, ifelse(fruit1<=199 & fruit1>=101,(fruit1-100)*30, 0))))
```
*Lets look at the mean, median, sd, and max for the new monthly values*
```{r}
fruits_and_vegetables %>%
  summarise("fruit mean" = mean(fruit_month, na.rm=TRUE), "median" = median(fruit_month, na.rm=TRUE), "standard deviation" = sd(fruit_month, na.rm=TRUE), "max" = max(fruit_month, na.rm=TRUE))
fruits_and_vegetables %>%
  summarise("veggies mean" = mean(veg_month, na.rm=TRUE), "median" = median(veg_month, na.rm=TRUE), "standard deviation" = sd(veg_month, na.rm=TRUE), "max" = max(veg_month, na.rm=TRUE))
```
*Since I don't think its possible to eat 2,970 of either frutis or vegetables we will look to remove some of the outliers. First lets use a skatter plot to view how the data is distributed, and check out the pearson correlation*
```{r}
 ggplot(data=fruits_and_vegetables, aes(x=fruit_month, y=veg_month, na.rm=TRUE)) +
   geom_point()
 cor(x=fruits_and_vegetables$fruit_month, y=fruits_and_vegetables$veg_month, method="pearson", use = "complete.obs" )
```

*Lets take the mean of each of the veg_month and fruits_month and add 3 sd*

1.  *veg_month = 23 + 3 X 24 = ***95** 

2.  *fruit_month = 30 + 3 X 30 = ***120**

*120 / 30 = 4 pieces of fruit a day. that seems like about as much as it would be possible to eat for our purposes I will increase that to 5 pieces of fruit(150) a day and 5 pieces of vegetables(150).  Given the mean and standard deviations for the data set this should include the vast majority of our data points while any number out side this interval is boarding on being an unrealistic amount of fruits and vegetables to eat in a typical month. Before modifying the table lets print out how many values we will be dropping*
```{r}
table(fruits_and_vegetables$fruit_month > 150)
table(fruits_and_vegetables$veg_month > 150)
```
*We will be removing 1069/(456908+1069) = 0.23% from the fruit respondents and  339/(452489+339) = 0.075% from the veggies*

*These numbers 0.23% and 0.075% are not the same as they represent values that are out side of 150/month.  there are respondents that put down 3000k pieces of fruit a month while suposodely eating a pedesrian 1-5 pieces of veggies a day both these values will be droped out as one of the tuple falls outside the 150/month threashold*
```{r}
 fruits_and_vegetables <- fruits_and_vegetables %>%
   mutate(fruit_month_ = ifelse(fruit_month<=150, fruit_month, NA)) %>%
   mutate(veg_month_ = ifelse(veg_month<=150, veg_month, NA))
```
*Lets plot these new values and check out their correlation*
```{r warning=FALSE}
 fruits_and_veggies_plot <- ggplot(data=fruits_and_vegetables, aes(x=fruit_month, y=veg_month, na.rm=TRUE)) +
   geom_point()
 fruits_and_veggies_plot + stat_density_2d(aes(fill = ..level..), geom="polygon")
 cor(x=fruits_and_vegetables$fruit_month, y=fruits_and_vegetables$veg_month, method="pearson", use="complete.obs")
```
*As we look at the two graphs the Pearson correlation coefficient went from 0.2962249 to 0.3157918 when we removed some of what I would call outliers.  A Pearson correlation of 0.3 or -.3 is considered weak, this is not what I expected.  I thought that there would be a stronger correlation between eating fruits and eating vegetables.  Let's move on and plot the average number of veggies and fruits eaten as compared to the participants reported over all health.*
```{r}

```

##### **Research quesion 2:** {#ResearchQuestion2}

*Find the index values for the desired columns*
```{r}
grep("smoke100", colnames(brfss2013))
grep("X_educag", colnames(brfss2013))
```
*Create a database with the selected columns*
```{r}
cols = c(68,263)
smoking_education <- brfss2013[,cols,drop=FALSE]
names(smoking_education) <- c("Smoke", "Education")
```
*Create Mosaic comparing smoking with the highest educational level completed*
```{r}
ggplot(data=smoking_education) +
  geom_mosaic(aes(x=product(Education, Smoke), fill=Education), na.rm=TRUE) +
  labs(x="values", title="Yes vs No")
```

*It aprears that attending college/technical school and grraduating high school have similar levels of smoking while graduating from college and not graduating from high school play a roll in wether you smoke or not lets view the data as a bar graph with the level of education as the independant varialbe*
*What are the percentages*
```{r}
smoking_edu_graph <- smoking_education %>%
  filter(!is.na(smoking_education$Education), !is.na(smoking_education$Smoke)) %>%
  group_by(Smoke, Education) %>%
  summarise(n = n())
ggplot(smoking_edu_graph, aes(x = Education, y = n, fill = Smoke)) + geom_bar(stat = "identity", position = position_dodge(0.6), alpha = 0.8) + scale_fill_brewer(palette = "Set1") + scale_y_continuous("count") + labs(title = "Smoking distribution by Education level") + coord_flip()
```

