---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages
for this project I will be useing ggplot and ggmosaic to display the data and dplyr for our dataframe manipulation

```{r load-packages, message = FALSE}
library(ggplot2)
library(ggmosaic)
library(dplyr)
```

### Load data


```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data

The Behavioral Risk Factor Surveilance System or BRFSS for short is a telephone survay conducted in all 50 states, the District of Columbia, Guam, and Puerto Rico.  The objective of the BRFSS is to "collect uniform, state-specific data on preventive health practices and risk behaviors that are linked to chronic diseases, injuries, and preventable infectious diseases that affect the adult population."  The data collected by the BRFSS is done through a collection of dialing landline and cell phone interviews that are randomized

* * *

## Part 2: Research questions

**Research quesion 1:**
Map out any relataionship between income and age represented in the data

**Research quesion 2:**
Explore the correlation between the number of cigaretts one reports smoking and if they wear a seatbelt while in a car.

**Research quesion 3:**
Explore the correlation between the number of fruits a subject eats a month verse the number of vegetables they eat.


* * *

## Part 3: Exploratory data analysis

**Research quesion 1:**
*Find the index values for the desired columns*
```{r}
grep("X_age80", colnames(brfss2013))
grep("income2", colnames(brfss2013))
```
*Create a database with the selected columns*
```{r}
cols = c(330,50)
income_age <- brfss2013[,cols,drop=FALSE]
```
*Mutate the columns to make the graph easier to read and drop the unnessessary*
```{r}
income_age <- income_age %>%
  mutate(age=X_age80) %>%
  mutate(income=ifelse(income2=="$75,000 or more", ">75K", ifelse(income2=="Less than $75,000", "<75K", ifelse(income2=="Less than $50,000", "<50K", ifelse(income2=="Less than $35,000", "<35K", ifelse(income2=="Less than $25,000", "<25K", ifelse(income2=="Less than $20,000", "<20K", ifelse(income2=="Less than $15,000", "<15K", ifelse(income2=="Less than $10,000", "<10K", NA)))))))))
```
*Map income by age showing quatiles*
```{r}
ggplot(data=income_age, aes(x=income, y=age)) +
  geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=2, notch=FALSE)
```

**Research quesion 2:**
*Find the index values for the desired columns*
```{r}
grep("smoke100", colnames(brfss2013))
grep("X_educag", colnames(brfss2013))
```
*Create a database with the selected columns*
```{r}
cols = c(68,263)
smoking_education <- brfss2013[,cols,drop=FALSE]
```
*Create Mosaic comparing smoking with the highest education one completed*
```{r}
ggplot(data=smoking_education) +
  geom_mosaic(aes(x=product(X_educag, smoke100), fill=X_educag), na.rm=TRUE) +
  labs(x="values", title="name of graph")
```
*What are the percentages*
```{r}
table(smoking_education$X_educag)/nrow(smoking_education)
```

**Research quesion 3:**

*Find the index values for the columns I wish to analysis*
```{r}
grep("vegetab1", colnames(brfss2013))
grep("fruit1", colnames(brfss2013))
```
*Create a data frame with just the vegetables and fruit columns*
```{r}
cols = c(82,78)
fruits_and_vegetables <- brfss2013[,cols, drop=FALSE]
```
*Create columns with mutated values that represent the number of monthy fruits and vegetables, approximate the number of weekly fruits and vegetables by mulitplying the value by 4 (4 weeks a month), and approximate the number of daily fruits and vegetables by mulitplying the daily value by 30 (30 days in a month)*
```{r}
fruits_and_vegetables <- fruits_and_vegetables %>%
  mutate(veg_month = ifelse(vegetab1<=399 & vegetab1 >=301, (vegetab1 - 300), ifelse(vegetab1<=299 & vegetab1>=201,(vegetab1-200)*4, ifelse(vegetab1<=199 & vegetab1>=101,(vegetab1-100)*30, 0)))) %>%
   mutate(fruit_month = ifelse(fruit1<=399 & fruit1 >=301, (fruit1 - 300), ifelse(fruit1<=299 & fruit1>=201,(fruit1-200)*4, ifelse(fruit1<=199 & fruit1>=101,(fruit1-100)*30, 0))))
```
*Lets look at the mean, median, sd, and max for the new monthly values*
```{r}
fruits_and_vegetables %>%
  summarise(mean_fruits = mean(fruit_month, na.rm=TRUE), median_fruits = median(fruit_month, na.rm=TRUE), sd_fruits = sd(fruit_month, na.rm=TRUE), max_fruits = max(fruit_month, na.rm=TRUE))
fruits_and_vegetables %>%
  summarise(mean_veg = mean(veg_month, na.rm=TRUE), median_veg = median(veg_month, na.rm=TRUE), sd_veg = sd(veg_month, na.rm=TRUE), max_veg = max(veg_month, na.rm=TRUE))
```
*Since I don't think its possible to eat 2,970 of either frutis or vegetables we will look to remove some of the outliers. First lets use a skatter plot to view how the data is distributed, and check out the pearson correlation*
```{r}
# ggplot(data=fruits_and_vegetables, aes(x=fruit_month, y=veg_month, na.rm=TRUE)) +
#   geom_point()
# cor(x=fruits_and_vegetables$fruit_month, y=fruits_and_vegetables$veg_month, method="pearson", use = "complete.obs" )
```
*Lets take the mean of each of the veg_month and fruits_month and add 2 sd and use that as our window of values to be used in my analysis*

1.  *veg_month = 23 + 2 X 24 = ***71** 

2.  *fruit_month = 30 + 2 X 30 = ***90**
```{r}
# fruits_and_vegetables <- fruits_and_vegetables %>%
#   mutate(fruit_month_ = ifelse(fruit_month<91, fruit_month, NA)) %>%
#   mutate(veg_month_ = ifelse(veg_month<72, veg_month, NA))
```
*Lets plot these new values and check out their correlation*
```{r}
# fruits_and_veggies_plot <- ggplot(data=fruits_and_vegetables, aes(x=fruit_month_, y=veg_month_, na.rm=TRUE)) +
#   geom_point()
# fruits_and_veggies_plot + stat_density_2d(aes(fill = ..level..), geom="polygon")
# cor(x=fruits_and_vegetables$fruit_month_, y=fruits_and_vegetables$veg_month_, method="pearson", use="complete.obs")
```
