---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(ggmosaic)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data


* * *

## Part 2: Research questions

**Research quesion 1:**
Explore the correlation between the number of fruits a subject eats a month verse the number of vegetables they eat.

**Research quesion 2:**
Explore the correlation between the number of cigaretts one reports smoking and if they wear a seatbelt while in a car.

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**
<br/>
*Find the index values for the columns I wish to analysis*
```{r}
grep("vegetab1", colnames(brfss2013))
grep("fruit1", colnames(brfss2013))
```
*Create a data frame with just the vegetables and fruit columns*
```{r}
cols = c(82,78)
fruits_and_vegetables <- brfss2013[,cols, drop=FALSE]
```
*Create columns with mutated values that represent the number of monthy fruits and vegetables, approximate the number of weekly fruits and vegetables by mulitplying the value by 4 (4 weeks a month), and approximate the number of daily fruits and vegetables by mulitplying the daily value by 30 (30 days in a month)*
```{r}
fruits_and_vegetables <- fruits_and_vegetables %>%
  mutate(veg_month = ifelse(vegetab1<=399 & vegetab1 >=301, (vegetab1 - 300), ifelse(vegetab1<=299 & vegetab1>=201,(vegetab1-200)*4, ifelse(vegetab1<=199 & vegetab1>=101,(vegetab1-100)*30, 0)))) %>%
   mutate(fruit_month = ifelse(fruit1<=399 & fruit1 >=301, (fruit1 - 300), ifelse(fruit1<=299 & fruit1>=201,(fruit1-200)*4, ifelse(fruit1<=199 & fruit1>=101,(fruit1-100)*30, 0))))
```
*Lets look at the mean, median, sd, and max for the new monthly values*
```{r}
mean(fruits_and_vegetables$veg_month, na.rm=TRUE)
median(fruits_and_vegetables$veg_month, na.rm=TRUE)
sd(fruits_and_vegetables$veg_month, na.rm=TRUE)
max(fruits_and_vegetables$veg_month, na.rm=TRUE)
mean(fruits_and_vegetables$fruit_month, na.rm=TRUE)
median(fruits_and_vegetables$fruit_month, na.rm=TRUE)
sd(fruits_and_vegetables$fruit_month, na.rm=TRUE)
max(fruits_and_vegetables$fruit_month, na.rm=TRUE)
```
*Since I don't think its possible to eat 2,970 of either frutis or vegetables we will look to remove some of the outliers. First lets use a skatter plot to view how the data is distributed, and check out the pearson correlation*
```{r}
# ggplot(data=fruits_and_vegetables, aes(x=fruit_month, y=veg_month, na.rm=TRUE)) +
#   geom_point()
# cor(x=fruits_and_vegetables$fruit_month, y=fruits_and_vegetables$veg_month, method="pearson", use = "complete.obs" )
```
*Lets take the mean of each of the veg_month and fruits_month and add 2 sd and use that as our window of values to be used in my analysis*<br/>
1.  *veg_month = 23 + 2 X 24 = ***71** <br/>
2.  *fruit_month = 30 + 2 X 30 = ***90**
```{r}
# fruits_and_vegetables <- fruits_and_vegetables %>%
#   mutate(fruit_month_ = ifelse(fruit_month<91, fruit_month, NA)) %>%
#   mutate(veg_month_ = ifelse(veg_month<72, veg_month, NA))
```
*Lets plot these new values and check out their correlation*
```{r}
# fruits_and_veggies_plot <- ggplot(data=fruits_and_vegetables, aes(x=fruit_month_, y=veg_month_, na.rm=TRUE)) +
#   geom_point()
# fruits_and_veggies_plot + stat_density_2d(aes(fill = ..level..), geom="polygon")
# cor(x=fruits_and_vegetables$fruit_month_, y=fruits_and_vegetables$veg_month_, method="pearson", use="complete.obs")
```
**Research quesion 2:**
*Find the index values for the desired columns*
```{r}
grep("smoke100", colnames(brfss2013))
grep("X_educag", colnames(brfss2013))
```
*Create a database with the selected columns*
```{r}
cols = c(68,263)
smoking_education <- brfss2013[,cols,drop=FALSE]
```
*Create Mosaic comparing smoking with the highest education one completed*
```{r}
ggplot(data=smoking_education) +
  geom_mosaic(aes(x=product(X_educag, smoke100), fill=X_educag), na.rm=TRUE) +
  labs(x="values", title="name of graph")
```


**Research quesion 3:**
*Find the index values for the desired columns*
```{r}
grep("X_age80", colnames(brfss2013))
grep("income2", colnames(brfss2013))
```
*Create a database with the selected columns*
```{r}
cols = c(330,50)
income_age <- brfss2013[,cols,drop=FALSE]
```
*Mutate the columns to make the graph easier to read and drop the unnessessary*
```{r}
income_age <- income_age %>%
  mutate(age=X_age80) %>%
  mutate(income=ifelse(income2=="$75,000 or more", ">75K", ifelse(income2=="Less than $75,000", "<75K", ifelse(income2=="Less than $50,000", "<50K", ifelse(income2=="Less than $35,000", "<35K", ifelse(income2=="Less than $25,000", "<25K", ifelse(income2=="Less than $20,000", "<20K", ifelse(income2=="Less than $15,000", "<15K", ifelse(income2=="Less than $10,000", "<10K", NA)))))))))
```
*Map income by age showing quatiles*
```{r}
ggplot(data=income_age, aes(x=income, y=age)) +
  geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=2, notch=FALSE)
```